Cognitive and Auditory Predictors of Dysarthric Speech Understanding in Older

### 2024-01-04 Manuscript Analyses/manuscript_analysis.Rmd
reorganizing rmd file

All data processing and analyses was conducted using R Version 4.3.2.

@@

@@

The overall demographic information for the participants are below.

@@

@@

In this study, listeners were assigned to one of two speakers for perceptual training: a speaker with Ataxic dysarthria (AM1) or a speaker with Parkinson's disease (PDM10). Means and standard deviation for listener age, cognition, and word in noise hearing threshold for each condition are presented below. Wilcoxon rank sum tests were performed for each of the measures. None of the cognition or hearing measures were significantly different between the two groups.

@@

pl_int %>%
  ggplot() +
  aes(x = type,
      y = pwc,
      group = id) +
  geom_line(alpha = 0.3) +
  stat_summary(aes(group = speaker),
               fun.y = mean, geom = "line", linewidth = 2) +
  facet_wrap("speaker") +
  labs(x = "Phase", y = "Intelligibility") +
  theme_bw() +
  theme(legend.position = "none", aspect.ratio = 1)

@@

AM1_cor <- cor_df %>%
  dplyr::filter(speaker == "AM1") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Picture Sequence Memory` = pic_seq_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c("#6D9EC1", "white", "#E46726"),
    lab = TRUE,
    insig = "blank",
    title = "Ataxic", 
    legend.title = "Correlation"
  )

@@

PDM10_cor <- cor_df %>%
  dplyr::filter(speaker == "PDM10") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Picture Sequence Memory` = pic_seq_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c("#6D9EC1", "white", "#E46726"),
    lab = TRUE,
    insig = "blank",
    title = "Parkinson's Disease", 
    legend.title = "Correlation"
  )

@@

rm(AM1_cor, PDM10_cor, cog_table, speaker_intel)

@@

@@

plot_fun = function(mods, title){
  plot_importance(
    mods = mods,
    labels = c("main", "int")
  ) %>% 
    drop_na(value) %>% 
    mutate(name = factor(
      name, 
      levels = c("main", "int"),
      labels = c(
        paste0("Main Effects"),
        paste0("Interactions")
    ))) %>% 
    filter(name == "Interactions") %>% 
    mutate(var = str_replace(var, "flanker_c", "Inhibitory Control of Attention"),
           var = str_replace(var, "card_sort_c", "Cognitive Flexibility"),
           var = str_replace(var, "vocab_c", "Vocabulary Knowledge"),
           var = str_replace(var, "pattern_c", "Processing Speed"),
           var = str_replace(var, "list_sort_c", "Working Memory"),
           var = str_replace(var, "pic_seq_c", "Picture Sequence Memory"),
           var = str_replace(var, "best_win", "Words in Noise Threshold"),
           var = str_replace(var, "_", " x ")) %>% 
    arrange(value) %>% 
    mutate(var = fct_inorder(var)) %>% 
    ggplot(aes(x = value, y = var, color = value)) +
    geom_segment(aes(xend = 0, yend = var)) +
    geom_point(size = 2) +
    facet_wrap(~name, scales = "free") +
    labs(x = "Relative Importance",
         y = "",
         title = {{title}}) +
    scale_color_viridis_c(option = "magma", end = .8) +
    theme(plot.title.position = "plot")
}

mods1 = mod_fun("AM1", "pretest")
mods2 = mod_fun("PDM10", "pretest")

p1 = plot_fun(mods1, "a) Ataxic")
p2 = plot_fun(mods2, "b) PD")

p2

@@

initial_mods %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

@@

t_for_diff(mods1)
t_for_diff(mods2)

@@

@@

mods12 = mod_fun2("AM1")
mods22 = mod_fun2("PDM10")

p12 = plot_fun(mods12, "a) Ataxic")
p22 = plot_fun(mods22, "b) PD")

p12
```

p22

@@

@@

t_for_diff(mods12)
t_for_diff(mods22)

@@

@@

>>>>

All data processing and analyses were conducted using R Version 4.3.2.

@@

First, pretest and posttest intelligibility was calculated as percent words correct (pwc) for each listener.

@@

A total of 60 older adult listeners participated in the study. The overall demographic information for the participants are below.

@@

# Removing unneeded items from the environment

rm(demo)

@@

In this study, listeners were assigned to one of two speakers for perceptual training: a speaker with Ataxic dysarthria (AM1) or a speaker with Parkinson's disease (PDM10). Means and standard deviation for listener age, cognition, and word in noise hearing threshold for each condition are presented below. Wilcoxon rank sum tests were performed for each of the measures. None of the cognition or hearing measures were significantly different between the two groups of listeners.

@@

pl_int %>%
  ggplot() +
  aes(x = type,
      y = pwc,
      group = id) +
  geom_line(alpha = 0.3) +
  stat_summary(aes(group = speaker),
               fun.y = mean, geom = "line", linewidth = 2) +
  facet_wrap("speaker") +
  labs(x = "Phase", y = "Intelligibility") +
  theme_classic() +
  theme(legend.position = "none", aspect.ratio = 1)

@@

# Customize plot colors

library(LaCroixColoR) #remotes::install_github("johannesbjork/LaCroixColoR")
my_pal <- lacroix_palette("PeachPear", n = 15, type = "continuous")
my_pal <- rev(my_pal)

# AM1 Cor Plot
AM1_cor <- cor_df %>%
  dplyr::filter(speaker == "AM1") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Picture Sequence Memory` = pic_seq_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c(my_pal[1], "white", my_pal[15]),
    lab = TRUE,
    sig.level = 0.05,
    insig = "blank",
    title = "Ataxic", 
    legend.title = "Correlation"
  )

@@

# PDM10 Cor Plot
PDM10_cor <- cor_df %>%
  dplyr::filter(speaker == "PDM10") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Picture Sequence Memory` = pic_seq_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c(my_pal[1], "white", my_pal[15]),
    lab = TRUE,
    insig = "blank",
    title = "Parkinson's Disease", 
    legend.title = "Correlation"
  )

@@

rm(AM1_cor, PDM10_cor, cog_table, speaker_intel, demo_table)

@@

### AM1 Models

@@

plot_fun = function(mods, title){
  plot_importance(
    mods = mods,
    labels = c("main", "int")
  ) %>% 
    drop_na(value) %>% 
    mutate(name = factor(
      name, 
      levels = c("main", "int"),
      labels = c(
        paste0("Main Effects"),
        paste0("Interactions")
    ))) %>% 
    filter(name == "Interactions") %>% 
    mutate(var = str_replace(var, "flanker_c", "Inhibitory Control of Attention"),
           var = str_replace(var, "card_sort_c", "Cognitive Flexibility"),
           var = str_replace(var, "vocab_c", "Vocabulary Knowledge"),
           var = str_replace(var, "pattern_c", "Processing Speed"),
           var = str_replace(var, "list_sort_c", "Working Memory"),
           var = str_replace(var, "pic_seq_c", "Picture Sequence Memory"),
           var = str_replace(var, "best_win", "Words in Noise Threshold"),
           var = str_replace(var, "_", " x ")) %>% 
    arrange(value) %>% 
    mutate(var = fct_inorder(var)) %>% 
    ggplot(aes(x = value, y = var, color = value)) +
    geom_segment(aes(xend = 0, yend = var)) +
    geom_point(size = 2) +
    facet_wrap(~name, scales = "free") +
    labs(x = "Relative Importance",
         y = "",
         title = {{title}}) +
    scale_color_gradientn(colors = my_pal) +
    theme_bw() +
    theme(plot.title.position = "plot")
}

mods1 = mod_fun("AM1", "pretest")
p1 = plot_fun(mods1, "a) Ataxic")
p1

@@

AM1_initial_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

@@

t_for_diff(mods1)

@@

### PDM10 Models

mods2 = mod_fun("PDM10", "pretest")
p2 = plot_fun(mods2, "b) PD")
p2

@@

mods12 = mod_fun2("AM1")
p12 = plot_fun(mods12, "a) Ataxic")
p12

@@

AM1_change_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

@@

t_for_diff(mods12)

@@

### PDM10 Models

mods22 = mod_fun2("PDM10")
p22 = plot_fun(mods22, "b) PD")
p22

PDM10_change_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

t_for_diff(mods22)

@@

mods_table %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")
